---
title: "Untitled"
output: html_document
date: "2024-01-12"
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(survival)
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(survminer)
library(ggfortify)
```

```{r}
survival_data <- read.csv(here("data", "clean_data", "kaplan_survival_data.csv")) %>% 
  select(!X) %>% 
  mutate(month = month(start_date),
         year = year(start_date))

survival_sum <- survival_data %>% 
  filter(species %in% c("sale", "saap", "same", "enca", "arca", "erfa")) %>% 
  mutate(alive = case_when(dead == 0 ~ 1,
                           dead == 1 ~ 0)) %>% 
  group_by(plot, end_date) %>% 
  reframe(plants_alive = sum(alive)) %>% 
  rename(start_date = end_date) 

par <- read.csv(here("data", "clean_data", "par_clean.csv")) %>% 
  select(!X) %>% 
  group_by(plot, date) %>% 
  reframe(par = mean(par, na.rm = T)) %>% 
  mutate(month = month(date),
         year = year(date))

soil_moisture <- read.csv(here("data", "clean_data", "soil_moisture_clean.csv")) %>% 
  select(!X) %>%
  group_by(plot, date) %>% 
  reframe(soil_moisture = mean(soil_moisture, na.rm = T)) %>% 
  mutate(month = month(date),
         year = year(date))

survival_environmental <- survival_data %>% 
  left_join(par, by = c("plot", "month", "year")) %>% 
  left_join(soil_moisture, by = c("plot", "month", "year")) %>% 
  mutate(community = as.factor(community)) %>% 
  left_join(survival_sum, by = c("plot", "start_date")) %>% 
  mutate(plants_alive = case_when(community %in% c(1,2) & start_time == 0 ~ 54,
                                  TRUE ~ plants_alive),
         plants_alive = replace_na(plants_alive, 0))

survival_environmental$community <- factor(survival_environmental$community, levels = c(3,2,1))

survival_environmental_chap <- survival_environmental %>%  
  filter(species %in% c("rhov", "adfa", "mala", "cecu", "ceol", "rhil", "hear")) %>% 
  mutate(chap_type = case_when(species %in% c("rhov", "mala", "adfa") ~ "facultive",
                               species %in% c("cecu", "ceol", "rhil", "hear") ~ "obligate"),
         chap_type2 = case_when(species %in% c("cecu", "ceol") ~ "obligate_seeder",
                               species %in% c("rhil", "hear") ~ "obligate_sprouter"))

```

Facilitation for facultive sprouters until early augst

```{r}
coxph(Surv(start_time, end_time, dead) ~ community, id = unique_id, data = survival_environmental_chap %>%  filter(chap_type == "facultive" & end_time < 204)) %>% tbl_regression(exp = T) ## 204 is through 8-5-2022
coxph(Surv(start_time, end_time, dead) ~ community, id = unique_id, data = survival_environmental_chap %>%  filter(chap_type == "facultive" & end_time > 204)) %>%  tbl_regression(exp = T)
```

No Facilitation for obligate species
```{r}
coxph(Surv(start_time, end_time, dead) ~ community, id = unique_id, data = survival_environmental_chap %>%  filter(chap_type == "obligate")) %>% 
  tbl_regression(exp = T)
```

Similarly, facilitation for all chaparral when looking at number of plants, but only until early august
```{r}
coxph(Surv(start_time, end_time, dead) ~ plants_alive, id = unique_id, data = survival_environmental_chap %>%  filter(end_time < 204)) %>%
  tbl_regression(exp =T)

coxph(Surv(start_time, end_time, dead) ~ plants_alive, id = unique_id, data = survival_environmental_chap %>%  filter(end_time > 204)) %>% 
  tbl_regression(exp = T)
```

Now showing facilitation seperated by chaparral type as a function of number of plants
```{r}
coxph(Surv(start_time, end_time, dead) ~ plants_alive + chap_type, id = unique_id, data = survival_environmental_chap %>% filter(end_time < 204)) %>% tbl_regression(exp = T)

coxph(Surv(start_time, end_time, dead) ~ plants_alive + chap_type, id = unique_id, data = survival_environmental_chap %>% filter(end_time > 204)) %>% tbl_regression(exp = T)

```

And Visualizing the above facilitation table

```{r}
facilitation_model <- coxph(Surv(start_time, end_time, dead) ~ plants_alive + chap_type, id = unique_id, data = survival_environmental_chap %>% filter(end_time < 204))

facilitation_data <- expand.grid(plants_alive = seq(from = 0, to = 54, by = 1),
                                 chap_type = c("obligate", "facultive")) %>% 
  mutate(start_time = 0, 
         end_time = 204,
         dead = 0) %>% 
  mutate(surv_prob = predict(facilitation_model, newdata = ., type = "survival") * 100,
         interval = predict(facilitation_model, newdata = ., type = "survival", se.fit = T, interval = "confidence")[[2]] * 100,
         upper = surv_prob + interval,
         lower = surv_prob - interval)

ggplot(facilitation_data, aes(x = plants_alive, y = surv_prob, col = chap_type))+
  geom_line()+
  geom_ribbon(aes(ymin =lower, ymax = upper), alpha = 0.5)+
  theme_classic()+
  ylim(0,100)
```

Showing how number of plants relates to soil moisture and par

```{r}
summary(glm(par~plants_alive, family = Gamma, data = survival_environmental %>% filter(end_time < 204 & par > 0.7)))
summary(glm(soil_moisture~plants_alive, data = survival_environmental%>% filter(end_time < 204)))

ggplot(survival_environmental%>% filter(end_time < 204 & par > 0.7), aes(x = plants_alive, y = par))+
  geom_point(aes(col = start_time))+
  geom_smooth(method = "glm",
              method.args ="Gamma")+
  theme_classic()

ggplot(survival_environmental%>% filter(end_time < 204), aes(x = plants_alive, y = soil_moisture))+
  geom_point()+
  geom_smooth(method = "glm")+
  theme_classic()

summary(glm(par~plants_alive, family = Gamma, data = survival_environmental %>% filter(end_time < 204 & community != 3)))
summary(glm(soil_moisture~plants_alive, data = survival_environmental%>% filter(end_time < 204 & community != 3)))

ggplot(survival_environmental%>% filter(end_time < 204 & community != 3), aes(x = plants_alive, y = par))+
  geom_point(aes(col = start_time))+
  geom_smooth(method = "glm",
              method.args ="Gamma")+
  theme_classic()

ggplot(survival_environmental%>% filter(end_time < 204 & community != 3), aes(x = plants_alive, y = soil_moisture))+
  geom_point()+
  geom_smooth(method = "glm")+
  theme_classic()
```

Now looking at survival including environmental variables with number of sage and chaparral type

```{r}
survival_environmental_chap_narrow <- survival_environmental_chap %>% filter(date.y != "2022-06-09") ##getting rid of june 2022 soil moisture, because it is funky and not believable. 

august_survival_environmental_chap <- survival_environmental_chap_narrow %>% 
  filter(end_time < 204) 
coxph(Surv(start_time, end_time, dead) ~ soil_moisture + par + plants_alive * chap_type, 
              id = unique_id, 
              data = august_survival_environmental_chap) %>% 
  tbl_regression(exp = T)


coxph(Surv(start_time, end_time, dead) ~ soil_moisture + par + plants_alive * chap_type, 
              id = unique_id, 
              data = survival_environmental_chap_narrow %>% filter(end_time >204)) %>% 
  tbl_regression(exp = T)
```

Visualizing above augst model

```{r}
full_august_model <- coxph(Surv(start_time, end_time, dead) ~ soil_moisture + par + plants_alive * chap_type, 
              id = unique_id, 
              data = august_survival_environmental_chap)

august_full_enviro_data_soil_moisture <- expand.grid("soil_moisture" = seq(from = min(august_survival_environmental_chap$soil_moisture, 
                                                                                      na.rm = T), 
                                                                           to = max(august_survival_environmental_chap$soil_moisture, 
                                                                                    na.rm = T), 
                                                                           by = 0.5),
                                                     "chap_type" = c("facultive", "obligate")) %>% 
  mutate(start_time = 0, 
         end_time = 203,
         dead = 0,
         par = median(august_survival_environmental_chap$par, na.rm = T),
         plants_alive = median(august_survival_environmental_chap$plants_alive)) %>% 
  mutate(prop_alive = predict(full_august_model, newdata = ., type = "surv") *100,
         interval = predict(full_august_model, newdata = ., type = "surv", se.fit = T, interval = "confidence")[[2]]*100,
         upper = prop_alive + interval,
         lower = prop_alive - interval)

ggplot(august_full_enviro_data_soil_moisture, aes(x = soil_moisture, y = prop_alive, col = chap_type))+
  geom_line()+
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5)+
  theme_classic()

august_full_enviro_data_par <- expand.grid("par" = seq(from = min(august_survival_environmental_chap$par, na.rm = T), 
                                                                           to = max(august_survival_environmental_chap$par, na.rm = T), 
                                                                           by = 0.01),
                                                     "chap_type" = c("facultive", "obligate")) %>% 
  mutate(start_time = 0, 
         end_time = 203,
         dead = 0,
         soil_moisture = median(august_survival_environmental_chap$soil_moisture, na.rm = T),
         plants_alive = median(august_survival_environmental_chap$plants_alive)) %>% 
  mutate(prop_alive = predict(full_august_model, newdata = ., type = "surv") *100,
         interval = predict(full_august_model, newdata = ., type = "surv", se.fit = T, interval = "confidence")[[2]]*100,
         upper = prop_alive + interval,
         lower = prop_alive - interval)

ggplot(august_full_enviro_data_par, aes(x = par, y = prop_alive, col = chap_type))+
  geom_line()+
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5)+
  theme_classic() 

august_full_enviro_data_plants <- expand.grid("plants_alive" = seq(from = min(august_survival_environmental_chap$plants_alive, na.rm = T), 
                                                                           to = max(august_survival_environmental_chap$plants_alive, na.rm = T), 
                                                                           by = 1),
                                                     "chap_type" = c("facultive", "obligate")) %>% 
  mutate(start_time = 0, 
         end_time = 203,
         dead = 0,
         soil_moisture = median(august_survival_environmental_chap$soil_moisture, na.rm = T),
         par = median(august_survival_environmental_chap$par, na.rm = T)) %>% 
  mutate(prop_alive = predict(full_august_model, newdata = ., type = "surv") *100,
         interval = predict(full_august_model, newdata = ., type = "surv", se.fit = T, interval = "confidence")[[2]]*100,
         upper = prop_alive + interval,
         lower = prop_alive - interval)

ggplot(august_full_enviro_data_plants, aes(x = plants_alive, y = prop_alive, col = chap_type))+
  geom_line()+
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5)+
  theme_classic() 
```

Visualizing above late summer model

```{r}
late_summer_chap <- survival_environmental_chap %>% filter(end_time > 204)

late_summer_full_model <- coxph(Surv(start_time, end_time, dead) ~ soil_moisture + par + plants_alive * chap_type, 
              id = unique_id, 
              data = late_summer_chap)

late_summer_full_enviro_data_soil_moisture <- expand.grid("soil_moisture" = seq(from = min(late_summer_chap$soil_moisture, 
                                                                                           na.rm = T),
                                                                                to = max(late_summer_chap$soil_moisture,
                                                                                         na.rm = T),
                                                                                by = 0.5),
                                                     "chap_type" = c("facultive", "obligate")) %>% 
  mutate(start_time = 0, 
         end_time = 419,
         dead = 0,
         par = median(late_summer_chap$par, na.rm = T),
         plants_alive = median(late_summer_chap$plants_alive)) %>% 
  mutate(prop_alive = predict(late_summer_full_model, newdata = ., type = "surv") *100,
         interval = predict(late_summer_full_model, newdata = ., type = "surv", se.fit = T, interval = "confidence")[[2]]*100,
         upper = prop_alive + interval,
         lower = prop_alive - interval)

ggplot(late_summer_full_enviro_data_soil_moisture, aes(x = soil_moisture, y = prop_alive, col = chap_type))+
  geom_line()+
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5)+
  theme_classic()

late_summer_full_enviro_data_par <- expand.grid("par" = seq(from = 0.6, 
                                                                           to = max(late_summer_chap$par, na.rm = T), 
                                                                           by = 0.01),
                                                     "chap_type" = c("facultive", "obligate")) %>% 
  mutate(start_time = 0, 
         end_time = 419,
         dead = 0,
         soil_moisture = median(late_summer_chap$soil_moisture, na.rm = T),
         plants_alive = median(late_summer_chap$plants_alive)) %>% 
  mutate(prop_alive = predict(late_summer_full_model, newdata = ., type = "surv") *100,
         interval = predict(late_summer_full_model, newdata = ., type = "surv", se.fit = T, interval = "confidence")[[2]]*100,
         upper = prop_alive + interval,
         lower = prop_alive - interval)

ggplot(late_summer_full_enviro_data_par, aes(x = par, y = prop_alive, col = chap_type))+
  geom_line()+
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5)+
  theme_classic() 

late_summer_full_enviro_data_plants <- expand.grid("plants_alive" = seq(from = min(late_summer_chap$plants_alive, na.rm = T), 
                                                                           to = max(late_summer_chap$plants_alive, na.rm = T), 
                                                                           by = 1),
                                                     "chap_type" = c("facultive", "obligate")) %>% 
  mutate(start_time = 0, 
         end_time = 419,
         dead = 0,
         soil_moisture = median(late_summer_chap$soil_moisture, na.rm = T),
         par = median(late_summer_chap$par, na.rm = T)) %>% 
  mutate(prop_alive = predict(late_summer_full_model, newdata = ., type = "surv") *100,
         interval = predict(late_summer_full_model, newdata = ., type = "surv", se.fit = T, interval = "confidence")[[2]]*100,
         upper = prop_alive + interval,
         lower = prop_alive - interval)

ggplot(late_summer_full_enviro_data_plants, aes(x = plants_alive, y = prop_alive, col = chap_type))+
  geom_line()+
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5)+
  theme_classic() 
```

